Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-10-06 paper/paper.tex
remove rug plot from Figure 3B as it is too small to be useful

\subsection{A system to identify and enrich for rare IFN+ positive cells}
A challenge in studying IFN induction by influenza virus is the rareness with which infection triggers its expression.
For instance, we previously performed single-cell transcriptomics on hundreds of influenza-infected A549 human lung epithelial cells at early times (6 to 10 hours) post-infection~\citep{russell2018extreme}, and found that only 1 of 368 infected cells expressed detectable type I or type III IFN (\FIG{IFNrare}A).
Prior studies using reporters have found that IFN expression is also rare in influenza-infected mice at 24 and even 48 hours post-infection~\citep{kallfass2013visualizing}, a conclusion supported by our own re-analysis of the single-cell transcriptomics of influenza-infected mice by \citet{steuerman2018dissection} (see \FIGSUPP[IFNrare]{mice}).

@@

\caption{
Single-cell transcriptomics of IFN-enriched influenza-infected cells.
{\bf (A)} 
Number of cells for which transcriptomes were obtained.
From these numbers, we estimate~\citep{bloom2018estimating} that $\approx$11\% of the transcriptomes are actually multiplets derived from multiple cells. 
{\bf (B)} The number of cellular and viral mRNAs detected for each cell is plotted as a point.
Green lines show the distribution of cellular mRNAs per cell, and the blue rug plot at the left of each panel shows the distribution of viral mRNAs per cell.
Cells outside the dashed magenta lines have unusually low or high amounts of cellular mRNA (likely low-quality emulsions or multiplets), and are excluded from subsequent analyses.
{\bf (C)} Distribution across cells of the fraction of all mRNA derived from influenza.
Cells called as infected are in blue, while other cells are in green.
The inset shows the amount of viral mRNA in the human cells that are called as infected.
{\bf (D)} Number of influenza genes detected per infected cell, and the amount of viral mRNA in cells expressing each number of viral genes.
The majority of cells express all eight gene segments, but a substantial minority fail to express at least one gene.
\FIGSUPP[transcriptomics]{frac_has_gene} shows the frequency that each viral gene is detected.
{\bf (E)} Relative expression of viral genes among infected cells.
{\bf (F)} Number of cells infected with wild-type virus, synonymously barcoded virus, or both.
From the cells infected with both viral barcodes, we estimate~\citep{bloom2018estimating} that 63\% of infected cells are co-infected.
{\bf (G)} Fraction of cellular mRNA from IFN across cells, faceted by whether the cells are infected.
Cells to the left of the first dashed magenta line are classified as IFN-, and cells to the right of the second line are classified as IFN+.
A pseudocount is added to the number of IFN transcripts detected in each cell, which is why none of the fractions are zero.
Many cells that do not express IFN still express ISGs (\FIGSUPP[transcriptomics]{ISG}).
}
\label{fig:transcriptomics}

>>>>

\subsection{A system to identify and enrich for rare IFN+ positive cells}
A challenge in studying IFN induction by influenza virus is the rareness with which infection triggers its expression.
We previously performed single-cell transcriptomics on hundreds of influenza-infected A549 human lung epithelial cells at early times (6 to 10 hours) post-infection~\citep{russell2018extreme}, and found that only 1 of 368 infected cells expressed detectable type I or type III IFN (\FIG{IFNrare}A).
Prior studies using reporters have found that IFN expression is also rare in influenza-infected mice at 24 and even 48 hours post-infection~\citep{kallfass2013visualizing}, a conclusion supported by our own re-analysis of the single-cell transcriptomics of influenza-infected mice by \citet{steuerman2018dissection} (see \FIGSUPP[IFNrare]{mice}).

@@

\caption{
Single-cell transcriptomics of IFN-enriched influenza-infected cells.
{\bf (A)} 
Number of cells for which transcriptomes were obtained.
From these numbers, we estimate~\citep{bloom2018estimating} that $\approx$11\% of the transcriptomes are actually multiplets derived from multiple cells. 
{\bf (B)} The number of cellular and viral mRNAs detected for each cell is plotted as a point.
Green lines show the distribution of cellular mRNAs per cell.
Cells outside the dashed magenta lines have unusually low or high amounts of cellular mRNA (likely low-quality emulsions or multiplets), and are excluded from subsequent analyses.
{\bf (C)} Distribution across cells of the fraction of all mRNA derived from influenza.
Cells called as infected are in blue, while other cells are in green.
The inset shows the amount of viral mRNA in the human cells that are called as infected.
{\bf (D)} Number of influenza genes detected per infected cell, and the amount of viral mRNA in cells expressing each number of viral genes.
The majority of cells express all eight gene segments, but a substantial minority fail to express at least one gene.
\FIGSUPP[transcriptomics]{frac_has_gene} shows the frequency that each viral gene is detected.
{\bf (E)} Relative expression of viral genes among infected cells.
{\bf (F)} Number of cells infected with wild-type virus, synonymously barcoded virus, or both.
From the cells infected with both viral barcodes, we estimate~\citep{bloom2018estimating} that 63\% of infected cells are co-infected.
{\bf (G)} Fraction of cellular mRNA from IFN across cells, faceted by whether the cells are infected.
Cells to the left of the first dashed magenta line are classified as IFN-, and cells to the right of the second line are classified as IFN+.
A pseudocount is added to the number of IFN transcripts detected in each cell, which is why none of the fractions are zero.
Many cells that do not express IFN still express ISGs (\FIGSUPP[transcriptomics]{ISG}).
}
\label{fig:transcriptomics}
